### 2025-04-05 code/processing-code/processed file.qmd

---
title: "Processing code"
author: "Murtaza Yaqubi"
format: html
editor: visual
---

# Load necessary libraries
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(here)           # For file paths
library(naniar)         # For gg_miss_var()
```

# PART 1: DATA IMPORT AND CLEANING
```{r}
# Import the dataset
cardio_df <- read_csv(here("data", "raw-data", "Cardiovascular_Disease_Dataset.csv"))

# Basic exploration of the dataset
head(cardio_df)           # View the first few rows
colnames(cardio_df)       # Check column names
dim(cardio_df)            # Check dimensions (should be 1025 rows x 14 columns)
summary(cardio_df)        # Get a summary of the data
glimpse(cardio_df)        # Glimpse at the data structure

# Visualize missing values for each variable
gg_miss_var(cardio_df)    # Display number of NAs

# Transform variables into suitable types and add descriptive labels
cardio_df <- cardio_df %>%
  mutate(
    gender = factor(gender, levels = c(0, 1), labels = c("Female", "Male")), 
    chestpain = factor(chestpain, levels = c(0, 1, 2, 3),
                       labels = c("Typical Angina", "Atypical Angina", "Non-anginal Pain", "Asymptomatic")),
    fastingbloodsugar = factor(fastingbloodsugar, levels = c(0, 1), labels = c("False", "True")),
    restingrelectro = factor(restingrelectro, levels = c(0, 1, 2),
                             labels = c("Normal", "ST-T Abnormality", "Left Ventricular Hypertrophy")),
    exerciseangia = factor(exerciseangia, levels = c(0, 1), labels = c("No", "Yes")),
    slope = factor(slope, levels = c(1, 2, 3), labels = c("Upsloping", "Flat", "Downsloping")),
    target = factor(target, levels = c(0, 1), labels = c("No Heart Disease", "Heart Disease"))
  )

# Print the transformed dataset for verification
print(cardio_df)
```

# PART 3: STATISTICAL ANALYSIS: DATA CLEANING, SUMMARIZATION, AND VISUALIZATION
```{r}
# Filter the data for individuals under 50 with normal cholesterol (<200 mg/dL) and create risk indicator variables for elevated blood pressure and ST depression.
cleaned_df <- cardio_df %>%
  filter(age < 50, serumcholestrol > 0, serumcholestrol < 200) %>%  
  mutate(
    # Create a binary variable for blood pressure:
    # "Elevated" if restingBP is between 120 and 129; otherwise, "Otherwise"
    elevated_BP = if_else(between(restingBP, 120, 129), "Elevated", "Otherwise"),

    # Create a binary indicator for ST depression:
    # "Elevated" if oldpeak is greater than 1; otherwise, "Normal"
    elevated_oldpeak = if_else(oldpeak > 1, "Elevated", "Normal")
  ) %>% 
  # Select the variables we want to keep
  select(age, gender, chestpain, restingBP, serumcholestrol, fastingbloodsugar, 
         maxheartrate, oldpeak, target, elevated_BP, elevated_oldpeak)

# Print a sample of the cleaned dataset
head(cleaned_df)

# Create a summary table by grouping the cleaned data by blood pressure status, ST depression status, and heart disease (target)
summary_table <- cleaned_df %>%
  group_by(elevated_BP, elevated_oldpeak, target) %>%
  summarise(count = n(), .groups = "drop")
print(summary_table)

summary_df = skimr::skim(summary_table)
print(summary_table)
# save to file
summarytable_file = here("results", "tables", "summarytable2.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Save the processed data
```{r}
save_data_location <- here::here("data/processed-data/cleaned_df.rds")
saveRDS(cleaned_df, file = save_data_location)
```