### 2025-04-17 code/eda-code/01_summary_stats.qmd

---
title: "Summary statistics"
author: "Murtaza"
format: html
editor: visual
---

```{r}
# Calculate summary statistics for selected variables
summary_stats <- cardio_df %>% 
  summarise(
    Age_min = min(age, na.rm = TRUE),
    Age_max = max(age, na.rm = TRUE),
    Age_mean = mean(age, na.rm = TRUE),
    Age_sd = sd(age, na.rm = TRUE),

    RestingBP_min = min(restingBP, na.rm = TRUE),
    RestingBP_max = max(restingBP, na.rm = TRUE),
    RestingBP_mean = mean(restingBP, na.rm = TRUE),
    RestingBP_sd = sd(restingBP, na.rm = TRUE),

    SerumCholestrol_min = min(serumcholestrol[serumcholestrol > 0], na.rm = TRUE),
    SerumCholestrol_max = max(serumcholestrol[serumcholestrol > 0], na.rm = TRUE),
    SerumCholestrol_mean = mean(serumcholestrol[serumcholestrol > 0], na.rm = TRUE),
    SerumCholestrol_sd = sd(serumcholestrol[serumcholestrol > 0], na.rm = TRUE),

    MaxHeartRate_min = min(maxheartrate, na.rm = TRUE),
    MaxHeartRate_max = max(maxheartrate, na.rm = TRUE),
    MaxHeartRate_mean = mean(maxheartrate, na.rm = TRUE),
    MaxHeartRate_sd = sd(maxheartrate, na.rm = TRUE),

    Oldpeak_min = min(oldpeak[oldpeak > 0], na.rm = TRUE),
    Oldpeak_max = max(oldpeak, na.rm = TRUE),
    Oldpeak_mean = mean(oldpeak, na.rm = TRUE),
    Oldpeak_sd = sd(oldpeak, na.rm = TRUE)
  )

# Reshape into long format, then wide table
summary_table_formatted <- summary_stats %>%
  pivot_longer(cols = everything(), names_to = "stat", values_to = "value") %>%
  separate(stat, into = c("Variable", "Statistic"), sep = "_(?=[^_]+$)") %>%
  pivot_wider(names_from = Variable, values_from = value)

# View formatted table
summary_table_formatted

# save to file
summarytable_file = here("results", "tables", "table1_summary_statistics.rds")
saveRDS(summary_table_formatted, file = summarytable_file)
```