### 2025-05-01 code/analysis-code/analysis-code.qmd
final touches

---
title: "Inferential Analysis"
author: "Murtaza Yaqubi"
format: html
editor: visual
---

## Load/install core libraries

```{r}
library(here)
library(readr)
library(dplyr)
library(tidyverse)
```

# Read in the data

```{r}
df_clean <- readRDS(here("data","processed-data","cleaned_df.rds"))
```

# Inferential Analysis

```{r}
# Numeric outcome
df_inf <- df_clean %>% 
  mutate(outcome_num = if_else(outcome == "Disease", 1L, 0L))

# Chi-square tests
chisq_bp <- chisq.test(table(df_inf$bp_status, df_inf$outcome))
chisq_st <- chisq.test(table(df_inf$st_status, df_inf$outcome))

# Save chi-square results to tables
chisq_bp_tbl <- broom::tidy(chisq_bp)
saveRDS(chisq_bp_tbl, here("results","tables","table2_chisq_bp.rds"))
chisq_st_tbl <- broom::tidy(chisq_st)
saveRDS(chisq_st_tbl, here("results","tables","table3_chisq_st.rds"))

# Print tests
print(chisq_bp)
print(chisq_st)

# Logistic base model
mod_base <- glm(outcome_num ~ bp_status + st_status, data=df_inf, family=binomial)
base_coef <- summary(mod_base)$coefficients
or_base <- exp(base_coef[ ,"Estimate"])
ci_base_l <- exp(base_coef[ ,"Estimate"] - 1.96*base_coef[ ,"Std. Error"])
ci_base_u <- exp(base_coef[ ,"Estimate"] + 1.96*base_coef[ ,"Std. Error"])
tbl_base_or <- tibble(Term = rownames(base_coef), OR = or_base, CI_low = ci_base_l, CI_up = ci_base_u)

# Save base-model ORs
saveRDS(tbl_base_or, here("results","tables","table4_or_base_model.rds"))
print(tbl_base_or)

# Interaction model
mod_int <- glm(outcome_num ~ bp_status*st_status, data=df_inf, family=binomial)
int_coef <- summary(mod_int)$coefficients
or_int <- exp(int_coef[ ,"Estimate"])
ci_int_l <- exp(int_coef[ ,"Estimate"] - 1.96*int_coef[ ,"Std. Error"])
ci_int_u <- exp(int_coef[ ,"Estimate"] + 1.96*int_coef[ ,"Std. Error"])
tbl_int_or <- tibble(Term = rownames(int_coef), OR = or_int, CI_low = ci_int_l, CI_up = ci_int_u)

# Save interaction-model ORs
saveRDS(tbl_int_or, here("results","tables","table5_or_interaction_model.rds"))
print(tbl_int_or)

# Adjusted model
mod_adj <- glm(outcome_num ~ bp_status*st_status + gender + fasting_blood_sugar, data=df_inf, family=binomial)
adj_coef <- summary(mod_adj)$coefficients
or_adj <- exp(adj_coef[ ,"Estimate"])
ci_adj_l <- exp(adj_coef[ ,"Estimate"] - 1.96*adj_coef[ ,"Std. Error"])
ci_adj_u <- exp(adj_coef[ ,"Estimate"] + 1.96*adj_coef[ ,"Std. Error"])
tbl_adj_or <- tibble(Term = rownames(adj_coef), OR = or_adj, CI_low = ci_adj_l, CI_up = ci_adj_u)

# Save adjusted-model ORs
saveRDS(tbl_adj_or, here("results","tables","table6_or_adjusted_model.rds"))
print(tbl_adj_or)

# ROC safeguard
preds_int <- predict(mod_int, type = "response")
if (length(unique(preds_int)) > 2 && length(unique(df_inf$outcome_num)) == 2) {
  auc_val <- auc(roc(df_inf$outcome_num, preds_int))
  print(paste("AUC:", round(auc_val, 3)))

  # Save AUC value for interaction model
  auc_tbl <- tibble(Model = "Interaction_Model", AUC = auc_val)
  saveRDS(auc_tbl, here("results", "tables", "table7_auc_interaction_model.rds"))
} else message("ROC not generated: single class or no variability")

# Recreate ROC object for interaction model
roc_int <- roc(df_inf$outcome_num, preds_int)

# Annotate and save ROC plot with AUC on-plot
ggroc(roc_int, legacy.axes = TRUE, color = "red") +
  labs(
    title = "ROC Curve for Interaction Model",
    x     = "False Positive Rate",
    y     = "True Positive Rate"
  ) +
  annotate(
    "text",
    x = 0.6, y = 0.2,
    label = paste("AUC =", round(auc_val, 3)),
    hjust = 0,
    size = 5,
    color = "black"
  ) +
  theme_wsj() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5)
  ) -> auc_plot_annotated

# Save annotated ROC plot
ggsave(
  filename = here("results", "figures", "fig9_roc_interaction_annotated.png"),
  plot     = auc_plot_annotated,
  width    = 10,
  height   = 5,
  units    = "in",
  dpi      = 300
)

plot(auc_plot_annotated)
```

# 6. Predictive Modeling

```{r}
# Prepare data for modeling
set.seed(123)
df_mod <- df_inf %>% 
  mutate(outcome = factor(outcome, levels=c("NoDisease","Disease")))

# Train-test split
set.seed(123)
data_split <- initial_split(df_mod, prop=0.7, strata=outcome)
train <- training(data_split); 
test <- testing(data_split)

# Recipe 
rec <- recipe(outcome ~ bp_status + st_status, data=train) %>% 
  step_dummy(all_nominal_predictors())
```

# Logistic regression model

```{r}
set.seed(123)
log_spec  <- logistic_reg() %>% set_engine("glm") %>% set_mode("classification")
log_wf    <- workflow() %>% add_model(log_spec) %>% add_recipe(rec)
log_fit   <- fit(log_wf, data = train)
log_preds <- predict(log_fit, test, type = "prob") %>% bind_cols(test)
log_auc   <- roc_auc(log_preds, truth = outcome, `.pred_Disease`)
log_acc   <- accuracy(predict(log_fit, test, type = "class") %>% bind_cols(test), truth = outcome, estimate = .pred_class)

# Save logistic modeling results
saveRDS(log_auc, here("results", "tables", "logistic_auc.rds"))
saveRDS(log_acc, here("results", "tables", "logistic_accuracy.rds"))
print(log_auc)
print(log_acc)

```

# LASSO logistic

```{r}
tune_ctrl  <- vfold_cv(train, v = 5)
set.seed(123)
lasso_spec <- logistic_reg(penalty = tune(), mixture = 1) %>% set_engine("glmnet") %>% set_mode("classification")
lasso_wf   <- workflow() %>% add_model(lasso_spec) %>% add_recipe(rec)
lasso_res  <- tune_grid(lasso_wf, resamples = tune_ctrl, grid = 20, metrics = metric_set(roc_auc))
best_lasso <- select_best(lasso_res, metric = "roc_auc")
lasso_final<- finalize_workflow(lasso_wf, best_lasso)
lasso_fit  <- fit(lasso_final, data = train)
lasso_preds<- predict(lasso_fit, test, type = "prob") %>% bind_cols(test)
lasso_auc  <- roc_auc(lasso_preds, truth = outcome, `.pred_Disease`)
lasso_acc  <- accuracy(predict(lasso_fit, test, type = "class") %>% bind_cols(test), truth = outcome, estimate = .pred_class)

# Save LASSO modeling results
saveRDS(lasso_auc, here("results", "tables", "lasso_auc.rds"))
saveRDS(lasso_acc, here("results", "tables", "lasso_accuracy.rds"))
print(lasso_auc)
print(lasso_acc)
```

# Random forest

```{r}
set.seed(123)
rf_spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 500) %>% set_engine("ranger") %>% set_mode("classification")
rf_wf   <- workflow() %>% add_model(rf_spec) %>% add_recipe(rec)
rf_res  <- tune_grid(rf_wf, resamples = tune_ctrl, grid = 20, metrics = metric_set(roc_auc))
best_rf <- select_best(rf_res, metric = "roc_auc")
rf_final<- finalize_workflow(rf_wf, best_rf)
rf_fit  <- fit(rf_final, data = train)
rf_preds<- predict(rf_fit, test, type = "prob") %>% bind_cols(test)
rf_auc  <- roc_auc(rf_preds, truth = outcome, `.pred_Disease`)
rf_acc  <- accuracy(predict(rf_fit, test, type = "class") %>% bind_cols(test), truth = outcome, estimate = .pred_class)

# Save Random Forest modeling results
saveRDS(rf_auc, here("results", "tables", "rf_auc.rds"))
saveRDS(rf_acc, here("results", "tables", "rf_accuracy.rds"))
print(rf_auc)
print(rf_acc)
```

# 7. Model Comparison

```{r}
model_comp <- tibble(
  Model    = c("Logistic", "LASSO", "Random Forest"),
  AUC      = c(log_auc$.estimate, lasso_auc$.estimate, rf_auc$.estimate),
  Accuracy = c(log_acc$.estimate, lasso_acc$.estimate, rf_acc$.estimate)
)
# Save model comparison table
saveRDS(model_comp, here("results", "tables", "model_comparison.rds"))
print(model_comp)
```