### 2025-05-01 code/eda-code/eda-code.qmd

---
title: "Exploratory Data Analysis"
author: "Murtaza"
format: html
editor: visual
---

# Load/install core libraries

```{r}
library(here)
library(readr)
library(dplyr)
library(tidyverse)
```

# Read in the data

```{r}
df_clean <- readRDS(here("data","processed-data","cleaned_df.rds"))
```

# Exploratory Data Analysis

```{r}
# Summary stats
summary_stats_tbl <- df_clean %>%
  summarise(
    Age_min           = min(age, na.rm = TRUE),
    Age_max           = max(age, na.rm = TRUE),
    Age_mean          = mean(age, na.rm = TRUE),
    Age_sd            = sd(age, na.rm = TRUE),
    RestingBP_min     = min(resting_bp, na.rm = TRUE),
    RestingBP_max     = max(resting_bp, na.rm = TRUE),
    RestingBP_mean    = mean(resting_bp, na.rm = TRUE),
    RestingBP_sd      = sd(resting_bp, na.rm = TRUE),
    Chol_min          = min(cholesterol, na.rm = TRUE),
    Chol_max          = max(cholesterol, na.rm = TRUE),
    Chol_mean         = mean(cholesterol, na.rm = TRUE),
    Chol_sd           = sd(cholesterol, na.rm = TRUE),
    HR_min            = min(max_heart_rate, na.rm = TRUE),
    HR_max            = max(max_heart_rate, na.rm = TRUE),
    HR_mean           = mean(max_heart_rate, na.rm = TRUE),
    HR_sd             = sd(max_heart_rate, na.rm = TRUE),
    ST_min            = min(st_depression, na.rm = TRUE),
    ST_max            = max(st_depression, na.rm = TRUE),
    ST_mean           = mean(st_depression, na.rm = TRUE),
    ST_sd             = sd(st_depression, na.rm = TRUE)
  ) %>%
  pivot_longer(everything(), names_to = "stat", values_to = "value") %>%
  separate(stat, into = c("Variable","Statistic"), sep = "_(?=[^_]+$)") %>%
  pivot_wider(names_from = Variable, values_from = value)

print(summary_stats_tbl)

# Save summary table
saveRDS(summary_stats_tbl, here("results","tables","table1_summary_statistics.rds"))

## Visualizations

# Correlation matrix
p_corr <- ggcorrplot(cor(df_clean %>% 
                 select(age, resting_bp, cholesterol, max_heart_rate, st_depression),
                         use = "pairwise.complete.obs"), lab=TRUE)
print(p_corr)
ggsave(filename = here("results","figures","fig1_correlation_matrix.png"), plot = p_corr, width = 10, height = 5)

# Histogram of Age 
p_age <- df_clean %>% 
  ggplot(aes(age)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black") + 
  geom_density(aes(y = after_stat(count) * 1.2), color = "red", size = 1.2) +
  labs(
    title = "Distribution of Age",
    x     = "Age (years)",
    y     = "Count"
  ) +
  theme_wsj() +
  theme(
    # re-enable & style the axis titles
    axis.title.x = element_text(size = 12, face = "bold", colour = "black"),
    axis.title.y = element_text(size = 12, face = "bold", colour = "black"),
    plot.title   = element_text(size = 18, face = "bold", hjust = 0.5)
  )
print(p_age)
ggsave(filename = here("results","figures","fig2_age_histogram.png"), plot = p_age, width = 10, height = 5)

# Histogram of Cholesterol 
p_chol <- df_clean %>% 
  ggplot(aes(cholesterol)) + 
  geom_histogram(bins = 30, fill = "gold", color = "black") + 
  geom_density(aes(y = after_stat(count) * 1.2), color = "red", size = 1.2) +
  labs(
    title = "Distribution of Cholesterol",
    x     = "Cholesterol Level (mg/dL)",
    y     = "Count"
  ) +
  theme_wsj() +
  theme(
    axis.title.x = element_text(size = 12, face = "bold", colour = "black"),
    axis.title.y = element_text(size = 12, face = "bold", colour = "black"),
    plot.title   = element_text(size = 18, face = "bold", hjust = 0.5)
  )
print(p_chol)
ggsave(filename = here("results","figures", "fig3_cholesterol_histogram.png"), plot = p_chol, width = 10, height = 5)

# Barplot of CVD prevalence by BP status
p_bp <- ggplot(df_clean, aes(x = bp_status, fill = outcome)) +
  geom_bar(position = "fill", width = 0.6) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "CVD Prevalence by Blood Pressure Status",
    x     = "Blood Pressure Status",
    y     = "Percent of Individuals",
    fill  = "CVD"
  ) +
  theme_wsj() +
  theme(
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    plot.title   = element_text(size = 18, face = "bold", hjust = 0.5)
  )
print(p_bp)
ggsave(filename = here("results","figures","fig4_bp_barplot.png"), plot = p_bp, width = 10, height = 5)

# Barplot of CVD prevalence by ST-Depression status
p_st <- ggplot(df_clean, aes(x = st_status, fill = outcome)) +
  geom_bar(position = "fill", width = 0.6) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "CVD Prevalence by ST-Depression Status",
       x = "ST-Depression Status", y = "Percent of Individuals", fill = "CVD") +
  theme_wsj() +
  theme(
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    plot.title   = element_text(size = 18, face = "bold", hjust = 0.5)
  )
print(p_st)
ggsave(filename = here("results","figures","fig5_st_barplot.png"), plot = p_st, width = 10, height = 5)

# Interaction Heatmap: counts by BP & ST-Depression, faceted by CVD
data_heat <- df_clean %>% count(bp_status, st_status, outcome)

p_heat <- ggplot(data_heat, aes(x = bp_status, y = st_status, fill = n)) +
  geom_tile(color = "white") +
  geom_text(aes(label = n), size = 4) +
  facet_wrap(~ outcome) +
  labs(
    title = "Counts by BP & ST Status (facet by Disease)",
    x = "BP Status",
    y = "ST Status",
    fill = "Count"
  ) +
  theme_wsj() +
  theme(
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    plot.title   = element_text(size = 15, face = "bold", hjust = 0.5)
  )
print(p_heat)
ggsave(
  filename = here("results", "figures", "fig6_interaction_heatmap.png"), plot     = p_heat, width    = 10, height   = 5)

# Boxplot of age distribution by CVD within BP categories
p_box <- ggplot(df_clean, aes(x = bp_status, y = age, fill = outcome)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Age Distribution by CVD Status within BP Categories",
       x = "Blood Pressure Status", y = "Age (years)", fill = "CVD") +
  theme_wsj() +
  theme(
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    plot.title   = element_text(size = 15, face = "bold", hjust = 0.5)
  )
print(p_box)
ggsave(filename = here("results","figures","fig7_box_age_bp.png"), plot = p_box, width = 10, height = 5)

# Scatterplot of ST-Depression vs. resting BP by CVD status
p_scatter <- ggplot(df_clean, aes(x = resting_bp, y = st_depression, color = outcome)) +
  geom_point(alpha = 1, size = 3) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(title = "ST-Depression vs. Resting BP by CVD Status",
       x = "Resting Blood Pressure (mmHg)", y = "ST-Depression (oldpeak)", color = "CVD") +
  theme_wsj() +
  theme(
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    plot.title   = element_text(size = 18, face = "bold", hjust = 0.5)
  )
print(p_scatter)
ggsave(filename = here("results","figures","fig8_scatter_st_bp.png"), plot = p_scatter, width = 10, height = 5)
```