### 2025-05-01 code/processing-code/processing-code.qmd

---
title: "Processing code"
author: "Murtaza Yaqubi"
format: html
editor: visual
---

# Load/instal core libraries

```{r}
library(readr)
library(here)
library(dplyr)
library(tidyverse)
```

# 2. Data Import & Initial Cleaning

```{r}
# Import raw data
df_raw <- read_csv(here("data", "raw-data", "Cardiovascular_Disease_Dataset.csv"))

# Rename columns for clarity
df <- df_raw %>%
  rename(
    patient_id         = patientid,
    age                = age,
    gender             = gender,
    chest_pain_type    = chestpain,
    resting_bp         = restingBP,
    cholesterol        = serumcholestrol,
    fasting_blood_sugar= fastingbloodsugar,
    resting_ecg        = restingrelectro,
    max_heart_rate     = maxheartrate,
    exercise_angina    = exerciseangia,
    st_depression      = oldpeak,
    slope              = slope,
    num_major_vessels  = noofmajorvessels,
    outcome            = target
  )

# Recode 0 cholesterol as NA and out-of-range (>564) to NA
df <- df %>%
  mutate(
    cholesterol = na_if(cholesterol, 0),
    cholesterol = if_else(cholesterol > 564, NA_real_, cholesterol)
  )

# Convert binary/nominal variables to factors with labels
df <- df %>%
  mutate(
    gender           = factor(gender, levels = c(0,1), labels = c("Female","Male")),
    chest_pain_type  = factor(chest_pain_type, levels = 0:3,
                              labels = c("Typical Angina","Atypical","Non-anginal","Asymptomatic")),
    fasting_blood_sugar = factor(fasting_blood_sugar, levels = c(0,1), labels = c("False","True")),
    resting_ecg      = factor(resting_ecg, levels=0:2,
                              labels=c("Normal","ST-T Abnormality","LV Hypertrophy")),
    exercise_angina  = factor(exercise_angina, levels=c(0,1), labels=c("No","Yes")),
    slope            = factor(slope, levels=1:3, labels=c("Upsloping","Flat","Downsloping")),
    outcome          = factor(outcome, levels=c(0,1), labels=c("NoDisease","Disease"))
  )
```

# 3. Study Population Filter

```{r}
# Focus: age < 50 & normal cholesterol (<200)
df_clean <- df %>%
  filter(age < 50, !is.na(cholesterol), cholesterol < 200) %>%
  mutate(
    bp_status = factor(if_else(between(resting_bp, 120, 129), "Elevated", "Normal")),
    st_status = factor(if_else(st_depression > 1, "Elevated", "Normal"))
  )

# Save cleaned data
saveRDS(df_clean, here("data","processed-data","cleaned_df.rds"))
```